---
title: "Merge"
author: "Jose Manuel"
date: "24 de mayo de 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercise 1

Create the dataframes to merge:

buildings <- data.frame(location=c(1, 2, 3), name=c("building1", "building2", "building3"))

data <- data.frame(survey=c(1,1,1,2,2,2), location=c(1,2,3,2,3,1), efficiency=c(51,64,70,71,80,58))

The dataframes, buildings and data have a common key variable called, “location”. Use the merge() function to merge the two dataframes by “location”, into a new dataframe, “buildingStats”.


```{r exercise1}

buildings <- data.frame(location=c(1, 2, 3), name=c("building1", "building2", "building3"))
data <- data.frame(survey=c(1,1,1,2,2,2), location=c(1,2,3,2,3,1), efficiency=c(51,64,70,71,80,58))

(buildingStats <- merge(buildings, data, by = "location"))

```


## Exercise 2

Give the dataframes different key variable names:

**buildings <- data.frame(location=c(1, 2, 3), name=c("building1", "building2", "building3"))**
**data <- data.frame(survey=c(1,1,1,2,2,2), LocationID=c(1,2,3,2,3,1), efficiency=c(51,64,70,71,80,58))**

The dataframes, buildings and data now have corresponding variables called, location, and LocationID. Use the merge() function to merge the columns of the two dataframes by the corresponding variables.


```{r exercise2}

buildings <- data.frame(location=c(1, 2, 3), name=c("building1", "building2", "building3"))
data <- data.frame(survey=c(1,1,1,2,2,2), LocationID=c(1,2,3,2,3,1), efficiency=c(51,64,70,71,80,58))

(buildingStats <- merge(buildings, data, by.x = "location", by.y = "LocationID"))

```


## Exercise 3

**Inner Join**:

The R merge() function automatically joins the frames by common variable names. In that case, demonstrate how you would perform the merge in Exercise 1 without specifying the key variable.


```{r exercise3}

buildings <- data.frame(location=c(1, 2, 3), name=c("building1", "building2", "building3"))
data <- data.frame(survey=c(1,1,1,2,2,2), location=c(1,2,3,2,3,1), efficiency=c(51,64,70,71,80,58))

(buildingStats <- merge(buildings, data))

```


## Exercise 4

**Outer Join**:

Merge the two dataframes from Exercise 1. Use the “all=” parameter in the merge() function to return all records from both tables. Also, merge with the key variable, “location”.

```{r exercise4}

buildings <- data.frame(location=c(1, 2, 3), name=c("building1", "building2", "building3"))
data <- data.frame(survey=c(1,1,1,2,2,2), location=c(1,2,3,2,3,1), efficiency=c(51,64,70,71,80,58))

(buildingStats <- merge(buildings, data, all = TRUE))

```


## Exercise 5

**Left Join**:

Merge the two dataframes from Exercise 1. Use the “all=” parameter in the merge() function to return all records from both tables. Also, merge with the key variable, “location”.

```{r exercise5}

buildings <- data.frame(location=c(1, 2, 3), name=c("building1", "building2", "building3"))
data <- data.frame(survey=c(1,1,1,2,2,2), location=c(1,2,3,2,3,1), efficiency=c(51,64,70,71,80,58))

(buildingStats <- merge(buildings, data, all.x = TRUE))

```


## Exercise 6

**Right Join**:

Merge the two dataframes from Exercise 1, and return all rows from the right table. Use the matching key from Exercise 1 to return matching rows from the left table.

```{r exercise6}

buildings <- data.frame(location=c(1, 2, 3), name=c("building1", "building2", "building3"))
data <- data.frame(survey=c(1,1,1,2,2,2), location=c(1,2,3,2,3,1), efficiency=c(51,64,70,71,80,58))

(buildingStats <- merge(buildings, data, all.y = TRUE))

```


## Exercise 7

**Cross Join**:

Merge the two dataframes from Exercise 1, into a “Cross Join” with each row of “buildings” matched to each row of “data”. What new column names are created in “buildingStats”?

```{r exercise7}

buildings <- data.frame(location=c(1, 2, 3), name=c("building1", "building2", "building3"))
data <- data.frame(survey=c(1,1,1,2,2,2), location=c(1,2,3,2,3,1), efficiency=c(51,64,70,71,80,58))

(buildingStats <- merge(buildings, data, by=NULL))

```


## Exercise 8

**Merging Dataframe rows**:

To join two data frames (datasets) vertically, use the rbind function. The two data frames must have the same variables, but they do not have to be in the same order.

```{r exercise8}

buildings <- data.frame(location=c(1, 2, 3), name=c("building1", "building2", "building3"))
buildings2 <- data.frame(location=c(5, 4, 6), name=c("building5", "building4", "building6"))

(buildingStats <- rbind(buildings, buildings2))

```


## Exercise 9:

To join two data frames (datasets) vertically, use the rbind function. The two data frames must have the same variables, but they do not have to be in the same order.

```{r exercise9}

buildings3 <- data.frame(location=c(7, 8, 9),
name=c("building7", "building8", "building9"),
startEfficiency=c(75,87,91))

(buildings3 <- buildings3[,-3])

```


## Exercise 10:

Instead of deleting the extra variables from buildings3. append the buildings, and buildings2 with the new variable in buildings3, (from Exercise 9). Set the new data in buildings and buildings2 , (from Exercise 8), to NA.

```{r exercise10}

buildings3 <- data.frame(location=c(7, 8, 9),
name=c("building7", "building8", "building9"),
startEfficiency=c(75,87,91))

buildings[ , "startEfficiency"] <- NA
buildings2[ , "startEfficiency"] <- NA

```

```{r exercise10a, echo=FALSE}

(buildings)

(buildings2)

```