---
title: "ggvis basics"
author: "Jose Manuel Valdeolmillos"
date: "11 de mayo de 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introducción a ggvis

Tutorial para el uso de ggvis. Se dividirá en 4 partes principales:

    - Dive into plotting with ggvis().
    - Add interactivity with the mouse and keyboard.
    - Create more types of graphic by controlling the layer type.
    - Build up rich graphics with multiple layers.
    
## Carga de datos
Usaremos CARS

```{r cars}

summary(cars)

```

## Inicio con visualizacion de graficos.

Se puede plotear asignando a variable o directamente:

```{r grafica1}

# Las variables se ponen con ~ delante
library(ggvis)

p <- ggvis(mtcars, x = ~wt, y = ~mpg)

# Elegimos un grafico de puntos
layer_points(p)

```

```{r grafica1b}

# Las variables se ponen con ~ delante
layer_points(ggvis(mtcars, x = ~wt, y = ~mpg))

```

### Usamos %>% para trabajar con ggvis y dplyr. Para facilitar.

```{r grafica2}

mtcars %>%
    ggvis(x = ~wt, y = ~mpg) %>%
    layer_points()

```


```{r grafica3}

library(dplyr)

# conversión de desplazamiento (disp) a litros.

mtcars %>%
    ggvis(x = ~mpg, y = ~disp) %>%
    mutate(disp = disp / 61.0237) %>%
    layer_points()

```

### Se puede añadir más variables, ej: **fill**, **stroke**, **size**, **shape**.

```{r grafica4}

mtcars %>%
    ggvis(x = ~mpg, y = ~disp, stroke = ~vs) %>%
    layer_points()

```

```{r grafica4b}

mtcars %>%
    ggvis(x = ~mpg, y = ~disp, fill = ~vs) %>%
    layer_points()

```

```{r grafica4c}

mtcars %>%
    ggvis(x = ~mpg, y = ~disp, size = ~vs) %>%
    layer_points()

```

```{r grafica4d}

mtcars %>%
    ggvis(x = ~mpg, y = ~disp, shape = ~factor(cyl)) %>%
    layer_points()

```

### Para cambiar los puntos, en vez de "=" se debe usar ":=". Asi:

```{r grafica5}

mtcars %>%
    ggvis(x = ~wt, y = ~mpg, fill := "red", stroke := "black") %>%
    layer_points()

```

```{r grafica5b}

mtcars %>%
    ggvis(x = ~wt, y = ~mpg, size := 300, opacity := 0.4) %>%
    layer_points()

```

```{r grafica5c}

mtcars %>%
    ggvis(x = ~wt, y = ~mpg, shape := "cross") %>%
    layer_points()

```

### Principalmente esto es lo mas usado y mas importante, se puede hacer mas...


## Interaccion (No funciona en Rmd, hay que trabajar desde codigo)

```{r interaccion1}

mtcars %>%
    ggvis(x = ~wt, y = ~mpg, 
          size := input_slider(10,100),
          opacity := input_slider(0,1)) %>%
    layer_points()

```

### Podemos usar otros tipos de gráficos.
Algunos van dentro del layer y en vez de ":=" se pone "="

```{r interaccion2}

mtcars %>%
    ggvis(x = ~wt) %>%
    layer_histograms(width = input_slider(0, 2, step = 0.10, label = "width"),
                     center = input_slider(0, 2, step = 0.05, label = "center"))

```

As well as **input_slider()**, ggvis provides **input_checkbox()**, **input_checkboxgroup()**, **input_numeric()**, **input_radiobuttons()**, **input_select()** and **input_text()**. See the examples in the documentation for how you might use each one.

You can also use keyboard controls with **left_right()** and **up_down()**. Press the left and right arrows to control the size of the points in the next example.

```{r interaccion3}

keys_s <- left_right(10, 1000, step = 50)
mtcars %>% ggvis(~wt, ~mpg, size := keys_s, opacity := 0.5) %>% layer_points()

```

Otro ejemplo. Aparece el tooltip:

```{r interaccion4}

mtcars %>% ggvis(~wt, ~mpg) %>% 
  layer_points() %>% 
  add_tooltip(function(df) df$wt)

```


## Layers

Existe la posibilidad de trabajar con 5 layers:

### Points
**layer_points()**, with properties **x**, **y**, **shape**, **stroke**, **fill**, **strokeOpacity**, **fillOpacity**, and **opacity**.

```{r layer1}

mtcars %>% ggvis(~wt, ~mpg) %>% layer_points()

```

### Paths and Polygons
**layer_paths()**

```{r layer2}

df <- data.frame(x = 1:10, y = runif(10))
df %>% ggvis(~x, ~y) %>% layer_paths()

```

Con **fill** trabajamos con poligonos

```{r layer2a}

t <- seq(0, 2 * pi, length = 100)
df <- data.frame(x = sin(t), y = cos(t))
df %>% ggvis(~x, ~y) %>% layer_paths(fill := "red")

```

### Filled Areas
Use properties **y** and **y2** to control the extent of the area.

```{r layer3}

df <- data.frame(x = 1:10, y = runif(10))
df %>% ggvis(~x, ~y) %>% layer_ribbons()

```

```{r layer3a}

df %>% ggvis(~x, ~y + 0.1, y2 = ~y - 0.1) %>% layer_ribbons()

```

### Rectangulos
**layer_rects()** The location and size of the rectangle is controlled by the **x**, **x2**, **y** and **y2** properties.

```{r layer4}

set.seed(1014)
df <- data.frame(x1 = runif(5), x2 = runif(5), y1 = runif(5), y2 = runif(5))
df %>% ggvis(~x1, ~y1, x2 = ~x2, y2 = ~y2, fillOpacity := 0.1) %>% layer_rects()

```

### Texto
**layer_text()** The text layer has many new options to control the apperance of the text: **text** (the label), **dx** and **dy** (margin in pixels between text and anchor point), **angle** (rotate the text), **font** (font name), **fontSize** (size in pixels), **fontWeight** (e.g. bold or normal), **fontStyle** (italic or normal)...

```{r layer5}

df <- data.frame(x = 3:1, y = c(1, 3, 2), label = c("AA", "BB", "CC"))
df %>% ggvis(~x, ~y, text := ~label) %>% layer_text()

```

Cambiando el tamaño

```{r layer5a}

df <- data.frame(x = 3:1, y = c(1, 3, 2), label = c("AA", "BB", "CC"))
df %>% ggvis(~x, ~y, text := ~label) %>% layer_text(fontSize := 80)

```

Cambio el angulo

```{r layer5b}

df <- data.frame(x = 3:1, y = c(1, 3, 2), label = c("AA", "BB", "CC"))
df %>% ggvis(~x, ~y, text := ~label) %>% layer_text(angle := 60, fontSize := 50)

```


## Compound Layers

Hay 4 principales...

### **layer_lines()** 
Which automatically orders by the x variable:

```{r Clayers1}

t <- seq(0, 2 * pi, length = 20)
df <- data.frame(x = sin(t), y = cos(t))
df %>% ggvis(~x, ~y) %>% layer_paths()

```

```{r Clayers1a}

df %>% ggvis(~x, ~y) %>% layer_lines()

```

**layer_lines()** is equivalent to **arrange()** + **layer_paths()**:

```{r Clayers1b}

df %>% ggvis(~x, ~y) %>% arrange(x) %>% layer_paths()

```

### **layer_smooths()** 
Fits a smooth model to the data, and displays predictions with a line. It’s used to highlight the trend in noisy data:

```{r Clayers2}

mtcars %>% ggvis(~wt, ~mpg) %>% layer_smooths()

```

```{r Clayers2a}

# Equivalente

smoothed <- mtcars %>% compute_smooth(mpg ~ wt)
smoothed %>% ggvis(~pred_, ~resp_) %>% layer_paths()

```

Podemos controlar el grado de con el parametro **span**

```{r Clayers2b}

span <- input_slider(0.2, 1, value = 0.75)
mtcars %>% ggvis(~wt, ~mpg) %>% layer_smooths(span = span)

```

### **Multiples layers**

```{r Clayers3}

mtcars %>% 
  ggvis(~wt, ~mpg) %>% 
  layer_smooths() %>% 
  layer_points()

```

```{r Clayer3a}

mtcars %>% ggvis(~wt, ~mpg) %>%
  layer_smooths(span = 1) %>%
  layer_smooths(span = 0.3, stroke := "red")

```


### **Layer_histograms()** and **layer_freqpolys()** 
Which allows you to explore the distribution of continuous. Both layers first bin the data with **compute_bin()** then display the results with either rects or lines.

```{r Clayer4}

mtcars %>% ggvis(~mpg) %>% layer_histograms()

```

```{r Clayer4a}

#Equivalente
binned <- mtcars %>% compute_bin(~mpg) 
binned %>% 
  ggvis(x = ~xmin_, x2 = ~xmax_, y2 = 0, y = ~count_, fill := "black") %>%
  layer_rects()

```


Mas cosillas en

<http://ggvis.rstudio.com/data-hierarchy.html>

<http://ggvis.rstudio.com/properties-scales.html>

<http://ggvis.rstudio.com/axes-legends.html>